---
title: "Data Science for Business with R"
author: "Course Book"
date: "`r Sys.Date()`"
output:
  pagedown::html_paged:
    toc: true
    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Document Limitations

The course DS4B 201-R: Data Science for Business with R by Matt Dancho has been reformatted into book format using RStudio's `pagedown` package.  (See https://pagedown.rbind.io for the full documentation.) This document is strictly limited to course attendees.  The copyright and distribution of this document are controlled.  The documented follows the same copyright and limited distributions as the primary course.  At no time may this documented be shared or maintained in a public or shared repository.

# Business Problem

Differentiating between productive and non-productive employees is difficult and highly subjective. Some employees are in the top 20% (all stars), some are in the middle 60% (productive), and some are in the bottom 20%. The bottom 20% can be OK or can be poor depending on the overall level of talent in your organization. Sometimes just because they are bottom 20% they are still good employees because the organization is such a high performing organization that everyone they employee is good. However, most organizations do have bad employees.

Now, in terms of defining productivity, if on a 1 to 4 scale with 1 being poor and 4 being the best that managers rate employees 3 to 4, you can usually be OK in viewing the 4’s as productive. They may not always be all stars, especially if 50% of people reviewed get 4’s but in reality only 20% should by the bell curve. However, if a manager rates someone as a 1, there is definitely a problem. These people need to either have immediate corrective actions taken or need to be cut loose. This is good attrition because the employee is a poor job/culture fit.

The analysis we are doing specifically looks at bad attrition, losing people that are 3’s and 4’s. I call them high performers, but in reality, I should be saying productive employees. This group is probably driving 80% of the business results.

Non-profits will need to be handled based on what the business finds important. While maximizing profit may not be the goal, there are still goals. For example, one goal might be to increase high school graduation rates in schools. This could be set up showing that the percentage of student dropouts are much higher in the district of focus and that increasing graduation rates is expected to reduce crime and increase workforce productivity. Crime reduction would have a savings associated with it. Workforce productivity would have a financial benefit associated with it. These are possible ways we can show financial value for a non-profit.

Non-Profit Cost Analysis: https://www.bridgespan.org/bridgespan/Images/articles/nonprofit-cost-analysis-toolkit/NonprofitCostsAnalysisToolkit.pdf

Non-Profit Strategic Planning: https://knowhownonprofit.org/organisation/strategy
The tools are quite similar to what you might find in a for-profit scenario. The maximization will be to maximize the financial delta between cost and benefit (to public) versus cost and revenue (for organization).

## What Happens When Good Employees Leave

*You takeaway our top 20 employees and we become a mediocre company* –Bill Gates

•	All employees turnover, some just faster
•	Impact of Attrition
  –	Varies based on employee productivity
  –	Can be measured
•	Bad turnover costs the organization



```{r, results='asis', echo = FALSE}
random_markdown = function(len = 100) {
  uri = knitr::image_uri(file.path(R.home('doc'), 'html', 'logo.jpg'))
  text = function(len) {
    trimws(paste(sample(c(letters, rep(' ', 10)), len, TRUE), collapse = ''))
  }
  id = function() paste(sample(letters, 8, TRUE), collapse = '')
  figure = function() {
    sprintf('![(#fig:%s)The R logo.](%s){width=%d%%}', id(), uri, sample(20:60, 1))
  }
  tab = paste(knitr::kable(head(mtcars[, 1:5])), collapse = '\n')
  table = function() {
    c(sprintf('Table: (#tab:%s)A table example.', id()), tab)
  }
  unlist(lapply(seq_len(len), function(i) {
    if (i %% 20 == 0) return(paste('#', text(sample(10:30, 1))))
    if (i %% 10 == 0) return(paste('##', text(sample(10:30, 1))))
    type = sample(1:3, 1, prob = c(.9, .03, .07))
    switch(type, text(sample(50:300, 1)), figure(), table())
  }))
}
cat(random_markdown(), sep = '\n\n')
```
